A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-23 latex/main.tex
Define deletion presense
Closes #886

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
Although uncommon relative to single nucleotide mutations,
deletions are an important component of
SARS-CoV-2 evolution \cite{Jeronimo2023},
contributing to the emergence of VOCs by enhancing,
e.g., efficiency for cell entry \cite{Cantoni2022} and
potential for antibody escape \cite{McCarthy2021}.
Using the ARG, we identified major deletion events
(which are propagated to many descendants)
as well as recurrent deletions
(which occur on at least two branches) in the ARG.
In Table~\ref{tab:major_dels},
we list the major deletion events in the ARG
which are each inherited by at least 10,000 descendant samples; and
in Table~\ref{tab:rec_dels},
we list the recurrent deletions in the ARG
which are each observed in at least 10,000 (0.4\%) samples
% [FIXME define how we measured this].
(i.e., for a deletion resulting from multiple events in the ARG,
this is the total samples inheriting the deletion).

>>>>

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
Although uncommon relative to single nucleotide mutations,
deletions are an important component of
SARS-CoV-2 evolution \cite{Jeronimo2023},
contributing to the emergence of VOCs by enhancing,
e.g., efficiency for cell entry \cite{Cantoni2022} and
potential for antibody escape \cite{McCarthy2021}.
Using the ARG, we identified major deletion events
(which are propagated to many descendants)
as well as recurrent deletions
(which occur on at least two branches) in the ARG.
In Table~\ref{tab:major_dels},
we list the major deletion events in the ARG
which are each inherited by at least 10,000 descendant samples; and
in Table~\ref{tab:rec_dels},
we list the recurrent deletions in the ARG
which are each observed in at least 10,000 (0.4\%) samples.
A deletion is said to be observed in a sample here 
if the gap character is present at the specified positions,
and the immediately adjacent positions are not the gap
character.
